---
title: "After Hours - How Financial Security Affects Student Success at School"
authors: "Adora Thienthanh Pham, Anika Indurkar, Ella Gabrielle Uchtmann, Richard Tran"
date: "11/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r source_files, echo = FALSE, message = FALSE, include = FALSE}
# Load the summary.R file, which itself sources all other
# chart files.
source("../source/summary.R")
```

```{r variable_initialization, echo = FALSE}
# Load charts into variables for ease of access.
# Create food_2016 dataframe.
food_df <- load_food_insecurity_data(2016)
food_insecurity_plot_2016 <- create_food_chart(food_df, 2016)

grad_rates <- graduation_rates
grad_rates_plot <- bar_chart(grad_rates)
```

## **ANALYSIS SUMMARY:** 

This is where our dynamic paragraph will go :)

## **PLOTS AND CHARTS**

**Chart 1:** Comparison of Relative Child Food Insecurity to Median Income in Washington

```{r chart1, echo = FALSE}
plot(food_insecurity_plot_2016)
```

This scatter plot illustrates the relation between **median household income** and **child food insecurity rates** across **Washington counties** for the year **2016**. The downward trend suggests that there exists some correlation between the continuous variables; that is, _as median income increases, the relative percentage of children facing food insecurity decreases for a given county_. Of particular note is the Whitman County data point: this county had one of the highest food insecurity rates and one of the lowest household median incomes out of all Washington counties in 2016. The plot further reinforces our initial hypothesis that indeed, median income household bears some influence on students' reliable access to food. Abiding by the principle that no student should have to worry about their next meal while attending school, these findings are significant in that they illustrate the potential economic struggles faced by parents in families federally recognized as low income.

**Chart 2:** 
```{r chart2, echo = FALSE}
plot(grad_rates_plot)
```