---
title: "After Hours - How Financial Security Affects Student Success at School"
authors: "Adora Thienthanh Pham, Anika Indurkar, Ella Gabrielle Uchtmann, Richard Tran"
date: "11/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

```{r source_files, echo = FALSE, message = FALSE, include = FALSE}
# Load the summary.R file, which itself sources all other
# chart files.
source("../source/summary.R")
source("../source/aggregate_table.R")
```

```{r variable_initialization, echo = FALSE}
# Load charts into variables for ease of access.
# Create food insecurity dataframe 2016 and generate plot.
food_df <- load_food_insecurity_data(2016)
food_insecurity_plot_2016 <- create_food_chart(food_df, 2016)

# Grab graduation rates dataframe and generate plot.
grad_rates <- graduation_rates
grad_rates_plot <- bar_chart(grad_rates)

# Grab race and poverty dataframe and generate plot.
whitman_county_data <- whitman_county_2016
race_poverty_plot <- pie_chart_race_and_poverty(whitman_county_data)

most_insecure_counties <- paste(pull(low_median_high_insecure, Geography),
                                sep = ", ")

lowest_grad_rate <- paste0(format(round(lowest_grad_rate * 100, 2), nsmall = 2), "%")
```

## **ANALYSIS SUMMARY:** 

The top five counties that have the lowest median incomes and highest percentages of food insecure children in Washington are `r most_insecure_counties`. Whitman county faces the highest percentage of food insecure children as well as lowest median income households. A factor about student’s identities that affects their socioeconomic status is their race. The racial group with the highest percentage of people below the poverty level in Whitman county was `r most_below_poverty_whitman` peoples. The racial group with the lowest percentage of people below the poverty level was `r least_below_poverty_whitman`. Even further, racial identity connects with graduation rates across Washington. The county with the highest graduation rate in Washington is `r county_highest_rate`. The lowest graduation rate in Washington is `r lowest_grad_rate`.

## **SUMMARY TABLE**

```{r summarytable, echo = FALSE}
kable(aggregate_data)
```
<br>

This aggregate table shows the correlation between race, county, household income and graduation rate.
The reason why this table is included is because it looks at the larger picture of our topic: How
Financial Security Affects Student Success at School. The table proves that intersectionality within
race, income level and the neighborhood that you live can predict -to some extent how well you will
be able to function in school. According to the table, white students are more likely to graduate 
and are more likely to have a high income compared to native american students. 



## **PLOTS AND CHARTS**

**Chart 1:** Comparison of Relative Child Food Insecurity Rates to Median Income in Washington

```{r chart1, echo = FALSE}
plot(food_insecurity_plot_2016)
```

This scatter plot illustrates the relation between **median household income and
child food insecurity rates across Washington counties in 2016**. The downward
trend suggests that there exists some correlation between the continuous variables;
that is, _as median income increases, the relative percentage of children facing
food insecurity decreases for a given county_. Of particular note is the Whitman
County data point: this county had one of the highest food insecurity rates and
one of the lowest household median incomes out of all Washington counties in
2016. The plot supports our initial hypothesis that indeed, median income
household bears some influence on students' reliable access to food. Abiding by
the principle that no student should have to worry about their next meal while
attending school, these findings are significant in that they illustrate the
potential economic struggles faced by parents in families federally recognized
as low income.

<br>

**Chart 2:** Bar Chart of Number of Students Graduating by Race in Washington State, 2019-2020

```{r chart2, echo = FALSE}
plot(grad_rates_plot)
```

The bar chart above represents the 2019-20 graduation rates of students by
counties in Washington. This chart was included to illustrate the correlation
between the graduation rate of students by county and their racial identity.
Understanding this correlation is important in order to determine what equitable
resources can be given to different racial groups. The chart concludes that
white students have the highest graduation rate and native Hawaiian/ Pacific
Islander being the lowest graduation rate.

<br>

**Chart 3:** Pie Chart of Racial Distributions for Peoples Identified as Being Below the Poverty Line in Whitman County, Washington, 2016

```{r chart3, echo = FALSE}
plot(race_poverty_plot)
```

This pie chart shows the percentage of racial groups that stand below the poverty
level in Whitman County, Washington. The data is based on data from 2016. The
section color relates to the race of the people. The section size relates to the
percentage of that race that are below the poverty level. This chart was included
to show the relationship between poverty and race in Washington counties to
understand equitable change that needs to be made for certain groups based on
past trends. The majority of people below the poverty level in Whitman County
are Alaskan and American Natives. The least percentage of people below the
poverty level are white and Pacific Islander peoples.